---
title: "Lab4"
author: "Bek"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    theme: spacelab
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task2

### Read in the data, copy and paste the last six lines of the data using tail()

```{r}

spruce.df= read.csv("SPRUCE.csv") #reading in the SPRUCE.csv data into spruce.df
spruce.tail = tail(spruce.df) 
spruce.tail #printing out last six lines of SPRUCE.csv data
```

## Task3 

### Load the library s20x and make a lowess smoother scatter plot (Height Vs BHDiameter) using trendscatter() (use f=0.5) record the plot.

```{r}
library(s20x) #load the library s20x 
trendscatter(Height~BHDiameter, f=0.5, data = spruce.df) #using trendscatter to print out HEIGHT vs BHDiameter
```


### o	Make a linear model object, spruce.lm=with(spruce.df,lm(Height~BHDiameter))

```{r}
spruce.lm=with(spruce.df,lm(Height~BHDiameter)) 
summary(spruce.lm)

```


###	Find the residuals using residuals(), put them into an object called height.res

```{r}
height.res = residuals(spruce.lm)
```


###	Find the fitted values using fitted() and place them in an object called height.fit.


```{r}
height.fit = fitted(spruce.lm)

```

###	Plot the residuals vs fitted values.

```{r}
plot(height.res~height.fit, bg = "BLUE", pch = 21, cex = 1.2)
```


###	Plot the residuals vs fitted values using trendscatter()

```{r}
trendscatter(height.res ~ height.fit)
```


### What shape is seen in the plot? Compare it with the curve made with the trendscatter function (second line after Task3). 









