---
title: "Lab4"
author: "Bek"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    theme: spacelab
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task2

### Read in the data, copy and paste the last six lines of the data using tail()

```{r}

spruce.df= read.csv("SPRUCE.csv") #reading in the SPRUCE.csv data into spruce.df
spruce.tail = tail(spruce.df) 
spruce.tail #printing out last six lines of SPRUCE.csv data
```

## Task3 

### Load the library s20x and make a lowess smoother scatter plot (Height Vs BHDiameter) using trendscatter() (use f=0.5) record the plot.

```{r}
library(s20x) #load the library s20x 
trendscatter(Height~BHDiameter, f=0.5, data = spruce.df) #using trendscatter to print out HEIGHT vs BHDiameter
```


### o	Make a linear model object, spruce.lm=with(spruce.df,lm(Height~BHDiameter))

```{r}
spruce.lm=with(spruce.df,lm(Height~BHDiameter)) #creating a linear model from HEIGHT~BHDiameter
summary(spruce.lm) #printing out the summary

```


###	Find the residuals using residuals(), put them into an object called height.res

```{r}
height.res = residuals(spruce.lm) #residuals() to assign the residuals to height.res object for further use
```


###	Find the fitted values using fitted() and place them in an object called height.fit.


```{r}
height.fit = fitted(spruce.lm) #fitted() method to assign fitted values to height.fit object for further use

```

###	Plot the residuals vs fitted values.

```{r}
plot(height.res~height.fit, bg = "BLUE", pch = 21, cex = 1.2) #uing plot method to plot residuals against fitted values, blue dot and 1.2 more size
```


###	Plot the residuals vs fitted values using trendscatter()

```{r}
trendscatter(height.res ~ height.fit)
```


### What shape is seen in the plot? Compare it with the curve made with the trendscatter function (second line after Task3). 

the shape resembles that of a quadratic function(parabola), it has a resemblance to the curve made previously for the first half


###	Using the plot() function and spruce.lm, make the residual plot.

```{r}
plot(spruce.lm, which =1)
```


### Check normality using the s20x function normcheck(). Please note that you may need to add an additional option to show the Shapiro-Wilk test (use ?normcheck )

```{r}
normcheck(spruce.lm,shapiro.wilk = TRUE)
```

### What is the pvalue for the Shapiro-Wilk test? What is the NULL hypothesis in this case?

the p value equals 0.29 which is greater than 0.05, meaning the null hypothesis cannot be rejected.

###	Write a sentence outlining your conclusions concerning the validity of applying the straight line to this data set.

Residuals vs fitted shows a quadratic fitted curve when plotted meanwhile it should have been in normal distribution with a mean of 0 ranging from -1 to 1. Therefore, straight line is not a good representation of the model.

## Task4

### Fit a quadratic to the points using the appropriate formula inside the lm() function and placing the output in the object quad.lm.

```{r}
quad.lm=lm(Height~BHDiameter + I(BHDiameter^2),data=spruce.df)
summary(quad.lm)

```


### Make a fresh scatter plot of Height Vs BHDiameter and add the quadratic curve to it.


```{r}
plot(Height~BHDiameter, bg = "Blue", data = spruce.df, pch =21)
myplot = function(x) {
  quad.lm$coef[1]+quad.lm$coef[2]*x+quad.lm$coef[3]*x^2
}

curve(myplot, lwd = 2, col ="Red", add = TRUE)
```


### Make quad.fit, a vector of fitted values and make a plot of the residuals vs fitted values, use plot() and quad.lm

```{r}
quad.fit = fitted(quad.lm)
plot(quad.lm, which = 1)

```


### Check normality

```{r}
normcheck(quad.lm,shapiro.wilk = TRUE)
```



###	What is the value of the p-value in the Shapiro-Wilk test? What do you conclude?



